Active Learning from Blackbox to Timed Connectors

### 2016-01-15 paper/paper.tex
.

\usepackage{amssymb}
\usepackage{cite}
\usepackage[normalem]{ulem}
% for a smile-face character
\usepackage{wasysym}
% listing for codes
\usepackage{listings}
% listing for golang
% NOTE listings-golang.sty is required
\usepackage{listings-golang}
% to support flow graphs
\usepackage{textcomp}

@@

% personal characters
\newcommand{\rblock}[0]{\circleddash}
\newcommand{\rread}[0]{\rhd}
\newcommand{\rnoread}[0]{\oslash}
\newcommand{\smap}[1]{[{#1}]}

@@

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=.4\textwidth]{./images/howto.png}
  \end{center}
  \caption{The Active Learning Framework}
  \label{fig:howto}
\end{figure}

@@

\begin{figure}[h]
  \begin{center}
    \input{./images/basic.tex}
  \end{center}
  \caption{Basic Reo Channels}
  \label{fig:basic}
\end{figure}

@@

\begin{figure}[h]
  \begin{center}
    \input{./images/nodes.tex}
  \end{center}
  \caption{Source, Sink and Mixed Nodes in Reo}
  \label{fig:node}
\end{figure}

With definition of more basic channels, it's easy to extend Reo to formalize coordination in
different areas. In this paper, we take timed Reo\cite{DBLP:conf/sefm/ArbabBBR04} as our formal
model. Timed Reo includes several timed channels, where the most commonly-used one, called
\emph{timer}, is also shown in Figure \ref{fig:basic}. A t-timer channel accepts any data item from
its source end and returns on its sink end a timeout signal after a delay of t time units.

@@

\begin{figure}[h]
  \begin{center}
    \input{./images/connectorexample.tex}
  \end{center}
  \caption{Coordination with Reo Connectors}
  \label{fig:reoconnector}
\end{figure}

@@

A Mealy machine is in some state $q\in S$ is always ready for inputs at any time. If an input symbol
$i\in I$ is provided, the Mealy machine generates output $\lambda(q,i)$ and jumps to state
$\delta(q,i)$.

@@

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=.5\textwidth]{./images/activelearning.png}
  \end{center}
  \caption{Active Automata Learning}
  \label{fig:activelearning}
\end{figure}

@@

\begin{example}[PMM Semantics of FIFO1 channel]
  \label{example:pmmfifo}
  The semantics of FIFO channel with source end $A$ and sink end $B$ can be defined as follows.
  \begin{itemize}
    \item[-] $S(\Sigma)=\{q_0\}\cup\{q_d|d\in\Sigma\}$
    \item[-] $I=\{A\}$, $O=\{B\}$, $s_0=q_0$
    \item[-] output function
      \begin{displaymath}
        \delta(\Sigma)(s,i)=\left\{
        \begin{array}[h]{ll}
          (B:\bot) & s=q_0\land i=(A:\_,B:\rnoread) \\
          (B:\bot) & s=q_d\land i=(A:\_,B:\rnoread) \\     
          (B:d) & s=q_0\land i=(A:d,B:\rread)\\
          (B:d) & s=q_d\land i=(A:\_,B:\rread)\\
          \rblock & s=q_d\land i=(A:d,B:\rnoread) \\
          \rblock & s=q_0\land i=(A:\bot,B:\rread) \\
        \end{array}
        \right.
      \end{displaymath}
    \item[-] transition function
      \begin{displaymath}
        \lambda(\Sigma)(s,i)=\left\{
        \begin{array}[h]{ll}
          q_d & s=q_0\land i=(A:d,B:\rnoread) \\
          q_d & s=q_{d'}\land i=(A:d,B:\rread) \\
          q_0 & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
  \end{itemize}
\end{example}

Besides, the concrete mealy machine (where $\Sigma=\{a\}$) is shown in Figure \ref{fig:pmmfifo}.
\begin{figure}[h]
  \begin{center}
    \input{./images/pmmfifo.tex}
  \end{center}
  \caption{PMM-based Semantics of $\smap{FIFO1}_\Sigma$, where $\Sigma=\{a\}$}
  \label{fig:pmmfifo}
\end{figure}

\begin{example}[PMM Semantics of Sync channel]
  The PMM-based semantics of a Sync channel with a source-end A and a sink-end B can be defined as:
  \begin{itemize}
    \item[-] $S(\Sigma)=\{q_0\}$, $I=\{A\}$, $O=\{B\}$, $s_0=q_0$
    \item[-] output function
      \begin{displaymath}
        \delta(\Sigma)(s,i)=\left\{
        \begin{array}[h]{ll}
          (B:\bot) & s=q_0\land i=(A:\bot,B:\rnoread) \\
          (B:d) & s=q_0\land i=(A:d,B:\rread)\\
          \rblock & s=q_0\land i=(A:\bot,B:\rread) \\
          \rblock & s=q_0\land i=(A:d,B:\rnoread) \\
        \end{array}
        \right.
      \end{displaymath}
    \item[-] transition function $\lambda(\Sigma)(s,i)=q_0$.
  \end{itemize}
\end{example}

\begin{example}[PMM Semantics of n-Timer channel]
  Considering a n-Timer channel with a source-end A and a sink-end B, we define its PMM-based
  semantics as:
  \begin{itemize}
    \item[-] $S(\Sigma)=\{q_{i,d}|1\leq i\leq n, d\in \Sigma\}\cup\{q_0\}$
    \item[-] $I=\{A\}$, $O=\{B\}$, $s_0=q_0$
    \item[-] output function
      \begin{displaymath}
        \delta(\Sigma)(s,i)=\left\{
        \begin{array}[h]{ll}
          (B:d) & s=q_{n,d}\land i=(A:d,B:\rread)\\
          \rblock & s=q_{n,d}\:\land \\
          & (i=T\lor i=(A:\_,B:\rnoread) \\
          \rblock & s=q_{j,d}\land 1 \leq j < n\land \\
          & (i=(A:\_,B:\rread)\lor \\
          & i=(A:d,B:\_))\\
          (B:\bot) & otherwise\\
        \end{array}
        \right.
      \end{displaymath}
    \item[-] transition function
      \begin{displaymath}
        \lambda(\Sigma)(s,i)=\left\{
        \begin{array}[h]{ll}
          q_{j+1,d} & s=q_{j,d}\land i=T\land0 < j < n\\
          q_0 & s=q_{n,d}\land i=(A:\bot,B:\rread) \\
          q_{1,d} & s=q_{n,d'}\land i=(A:d,B:\rread) \\
          q_{1,d} & s=q_0 \land i=(A:d,B:\_) \\
          s & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
  \end{itemize}
\end{example}

@@

\todo{check the timed behavior}
\begin{definition}[Production Operator]
  Now we're going to define the production operator \emph{prod} of two parameterized Mealy machines as,
  \[
  prod(PM_1,PM_2)=PM_3
  \]
  as follows. Here we assume that $PM_2.O\cap PM_1.I=\varnothing$
  \begin{itemize}
  	\item[-] $\forall\Sigma, PM_3.S(\Sigma)=PM_1.S(\Sigma)\times PM_2.S(\Sigma)$
    \item[-] $PM_3.I=PM_1.I\cup PM_2.I-PM_1.O$
    \item[-] $PM_3.O=PM_1.O\cup PM_2.O-PM_2.I$
  \end{itemize}
  \emph{(Here we assume that sink ends of $PM_1$ can be connected to source ends $PM_2$, but not
  vise versa)}
  \begin{itemize}
    \item[-] $PM_3.s_0=(PM_1.s_0, PM_2.s_0)$
    \item[-] $\forall\Sigma, PM_3.\delta(\Sigma)((s_1,s_2), i)=$
      \begin{displaymath}
        \left\{
        \begin{array}[h]{ll}
          (Out_1 + Out_2)|_{PM_3.O} & Out_1\neq\rblock\land Out_2\neq\rblock \\
          \rblock & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
      where we have
      \begin{itemize}
        \item[*] $In_1 = i|_{PM_1.I}$
        \item[*] $Out_1 = PM_1.\delta(\Sigma)(s_1,In_1)$
        \item[*] $In_2 = (Out_1 + i)|_{PM_2.I}$
        \item[*] $Out_2 = PM_2.\delta(\Sigma)(s_2,In_2)$
      \end{itemize}
    \item[-] $\forall\Sigma, PM_3. \lambda(\Sigma)((s_1,s_2),i)=(s_1',s_2')$
      where we have
      \begin{itemize}
        \item[*] $s_1' = PM_1.\lambda(\Sigma)(s_1,In_1)$
        \item[*] $s_2' = PM_2.\lambda(\Sigma)(s_2,In_2)$
      \end{itemize}
  \end{itemize}
\end{definition}

@@

  \begin{itemize}
  	\item[-] $\forall\Sigma, PM'.S(\Sigma)=PM.S(\Sigma)$
    \item[-] $PM'.I=PM.I-\{IN\}$
    \item[-] $PM'.O=PM.O-\{OUT\}$
    \item[-] $PM'.s=PM.s$
    \item[-] $\forall\Sigma, i\in Input(\Sigma,PM'.I,PM'.O)$, $PM'.\delta(\Sigma)(s, i)=$
      \begin{equation}
        \label{equ:existd}
        PM.\delta(\Sigma)(s,T)|_{PM'.O}
      \end{equation}
      if $i=T$. When $i$ is a non-time action, we need to check whether there's a data item
      $d\in\Sigma$
      satiefies
      \[
      PM.\delta(\Sigma)(s,i\cup\{IN:d\})|_{OUT} = d
      \]
      if \emph{true}, we have $PM'.\delta(\Sigma)(s, i)=$
      \[
      PM.\delta(\Sigma)(s,i\cup\{IN:d\})|_{PM'.O}
      \]
      otherwise $PM'.\delta(\Sigma)(s, i)=\rblock$
    \item[-] $\forall\Sigma, i\in Input(\Sigma,PM'.I,PM'.O).'$ the transition
      function can be defined as $PM.\lambda(\Sigma)(s,i)=$
      \begin{displaymath}
        \left\{
        \begin{array}[h]{ll}
          PM.\lambda(\Sigma,T) & i=T \\
          PM.\lambda(\Sigma,i\cup\{IN:d\}) & \exists d\in\Sigma \mbox{ satisfes Equation
          \ref{equ:existd}} \\
          s & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
  \end{itemize}
\end{definition}

@@

\section{From Blackbox to Timed Connectors} 
\label{sec:activelearning}
In this section, we will show how to use L* algorithm to extract models from blackboxes, which
mainly includes 3 steps: \emph{constructing hypothesis}, \emph{enclosing hypothesis} and
\emph{checking hypothesis}.

\subsection{Observation Table}

\subsection{Closeness}

\subsection{Counter-Examples}
Generally, equivalence query has been proved impossible in blackbox
models\cite{DBLP:journals/iandc/Angluin87}. However, in this
section, we're showing that in certain circumstances, equivalence query can be implemented with
no approximation.

Equivalence queries are used to search for counter-examples. But what makes counter-examples even
existing? In Reo models, we believe that \emph{FIFO} channels and \emph{Timer} channels are to
blame. Here we present a brief example. To make things eaiser, we have only untimed Reo here.

\begin{figure}[h]
  \begin{center}
    \input{./images/3buf.tex}
  \end{center}
  \caption{A Switching Connector $S$ with Three Buffers}
  \label{fig:buf3}
\end{figure}

A \emph{Switching} connector in \figurename \ref{fig:buf3} has two source-ends $A,B$ and one
sink-end $C$. In a nutshell, datum come from $A$ and be stored temporarily in buffers. These datum
will never flow out until signals come to $B$. With the Mealy-Machine semantics given, the semantics
of \emph{Switching} connectors can be defined as following \figurename \ref{fig:buf3semantics}.

\begin{figure}[h]
  \begin{center}
    \input{./images/3bufsemantics.tex}
  \end{center}
  \caption{\emph{Gate} as Mealy Machine $\mathcal{M}(S)$}
  \label{fig:buf3semantics}
\end{figure}

According to the production of Mealy-Machines, 8 different states should be found in
$\mathcal{M}(S)$. We use $q_{abc}$ to denote them. $q_{001}$ indicates that the last buffer is
filled and others are empty, and $q_{111}$ means that there are no space in any buffer. It's
obvious that some states like $q_{100}$ is unreachable.

We say if two states are \emph{similiar} iff. they have the same output signature.

\begin{theorem}[Bound of Counter-Example]
  \label{the:cebound} Assume that an observation table has already been closed with maximum
  query length \todo{find a better phrase?} $l$. An counter-example of length $l+2$ would be
  found iff. there're possible counter-examples.
\end{theorem}
\begin{IEEEproof}
  Sketch of this proof are shown in several points:
  \begin{enumerate}
    \item Hypothesises are subgraphs of the semantics graph.
    \item Inputs on reverse edges are complementary.
    \item 
  \end{enumerate}
\end{IEEEproof}

@@

\subsection{Case Studies}
A simple example of timed connector is presented to show how L* works.
\begin{figure}[h]
  \begin{center}
    \input{./images/expfifo.tex}
  \end{center}
  \caption{Expiring FIFO1 Channel (ExpFIFO1)}
  \label{fig:expfifo}
\end{figure}

@@

\begin{table}[h]
  \renewcommand{\arraystretch}{1.3}
  \caption{Time-Cost Analysis}
  \label{tabel:timecost}
  \centering
  \begin{tabular}{l||rrrr}
    \hline
    & FIFO & Alternator & Gate & ExpFIFO1\\
    \hline\hline
    Membership Query(s) & 41.571 & 126.468 & 169.161 &\\
    Hypothesis Query(s) & 0.001 & 0.003 & 0.004 &\\
    Total Time(s) & 41.715 & 165.114 & 247.098 &\\
    Membership Query(\%) & 99.6 & 76.6 & 68.5 &\\
    \hline
  \end{tabular}
\end{table}

@@

If previous simulation results are stored in a well-maintained cache, the time-cost in
simulation process could be reduced signficiantly. In this work, we use a multiway tree to buffer
these results. A brief example of such trees can be found in Figure \ref{fig:multiway}.

\todo{the graph need further details}
\begin{example}[Multiway-Tree Cache]
  \label{example:tree}
  Considering the FIFO channel presented in Example \ref{example:pmmfifo}. 
  \begin{figure}[h]
    \begin{center}
      \input{./images/cache.tex}
    \end{center}
    \caption{Multiway-Tree Cache}
    \label{fig:multiway}
  \end{figure}
  Note that in this figure, only the edge information is stored. States like $q_0,q_a$ are written
  here only to make it clear.
\end{example}

\begin{table}[h]
  \renewcommand{\arraystretch}{1.3}
  \caption{Reduction of Membership Queries}
  \label{tabel:cacheoptimization}
  \centering
  \begin{tabular}{l||rrr}
    \hline
    & FIFO & Alternator & Gate \\
    \hline\hline
    Original Algorithm & 93 & 880 & 1034 \\
    Cached Algorithm & 90 & 725 & 707 \\
    Reduction Rate & 3.2\% & 21.4\% & 31.6\% \\
    \hline
  \end{tabular}
\end{table}

With cache applied, we have made a good reduction on the calls of membership queries. The results
can be found in Table \ref{tabel:cacheoptimization}.

>>>>

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{cite}
\usepackage[normalem]{ulem}
% for a smile-face character
\usepackage{wasysym}
% listing for codes
\usepackage{listings}
% listing for golang
% NOTE listings-golang.sty is required
\usepackage{listings-golang}
% to support flow graphs
\usepackage{textcomp}
% highlight in tables
\usepackage{xcolor}
\usepackage{colortbl}

\usepackage[linesnumbered, ruled, vlined]{algorithm2e}

@@

% personal characters
\newcommand{\rblock}[0]{\circleddash}
\newcommand{\rread}[0]{\rhd}
\newcommand{\rnoread}[0]{\oslash}
\newcommand{\smap}[1]{[{#1}]}
\newcommand{\rempty}[0]{\varnothing}

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=.4\textwidth]{./images/howto.png}
  \end{center}
  \caption{The Active Learning Framework}
  \label{fig:howto}
\end{figure}

@@

\begin{figure}[ht]
  \begin{center}
    \input{./images/basic.tex}
  \end{center}
  \caption{Basic Reo Channels}
  \label{fig:basic}
\end{figure}

@@

\begin{figure}[ht]
  \begin{center}
    \input{./images/nodes.tex}
  \end{center}
  \caption{Source, Sink and Mixed Nodes in Reo}
  \label{fig:node}
\end{figure}

With definition of more basic channels, it's easy to extend Reo to formalize coordination in
different areas. In this paper, we take timed Reo\cite{DBLP:conf/sefm/ArbabBBR04} as our formal
model. Timed Reo includes several timed channels, where the most commonly-used one, called
\emph{timer}, is also shown in Figure \ref{fig:basic}. A t-timer channel accepts any data item from
its source end, and later dispense it to its sink end after a delay of t time units.

@@

\begin{figure}[ht]
  \begin{center}
    \input{./images/connectorexample.tex}
  \end{center}
  \caption{Coordination with Reo Connectors}
  \label{fig:reoconnector}
\end{figure}

@@

A Mealy machine is in some state $q\in S$ is always ready for inputs at any time. If an input symbol
$i\in I$ is provided, the Mealy machine generates output $o=\lambda(q,i)$ and jumps to state
$q'=\delta(q,i)$. Such an edge is also denoted as $q\xrightarrow[]{i/o}q'$.

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=.5\textwidth]{./images/activelearning.png}
  \end{center}
  \caption{Active Automata Learning}
  \label{fig:activelearning}
\end{figure}

@@

\begin{example}[PMM Semantics of FIFO1 channel]
  \label{example:pmmfifo}
  The semantics of FIFO channel with source end $A$ and sink end $B$ can be defined as follows.
  \begin{itemize}
    \item[-] $S(\Sigma)=\{q_0\}\cup\{q_d|d\in\Sigma\}$
    \item[-] $I=\{A\}$, $O=\{B\}$, $s_0=q_0$
    \item[-] output function
      \begin{displaymath}
        \delta(\Sigma)(s,i)=\left\{
        \begin{array}[ht]{ll}
          (B:\bot) & s=q_0\land i=(A:\_,B:\rnoread) \\
          (B:\bot) & s=q_d\land i=(A:\_,B:\rnoread) \\     
          (B:d) & s=q_0\land i=(A:d,B:\rread)\\
          (B:d) & s=q_d\land i=(A:\_,B:\rread)\\
          \rblock & s=q_d\land i=(A:d,B:\rnoread) \\
          \rblock & s=q_0\land i=(A:\bot,B:\rread) \\
        \end{array}
        \right.
      \end{displaymath}
    \item[-] transition function
      \begin{displaymath}
        \lambda(\Sigma)(s,i)=\left\{
        \begin{array}[ht]{ll}
          q_d & s=q_0\land i=(A:d,B:\rnoread) \\
          q_d & s=q_{d'}\land i=(A:d,B:\rread) \\
          q_0 & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
  \end{itemize}
\end{example}

Besides, the concrete mealy machine (where $\Sigma=\{a\}$) is shown in Figure \ref{fig:pmmfifo},
where labels of edges are in form of $\frac{input}{output}$.
\begin{figure}[ht]
  \begin{center}
    \input{./images/pmmfifo.tex}
  \end{center}
  \caption{PMM-based Semantics of $\smap{FIFO1}_\Sigma$, where $\Sigma=\{a\}$}
  \label{fig:pmmfifo}
\end{figure}

\begin{example}[PMM Semantics of Sync channel]
  The PMM-based semantics of a Sync channel with a source-end A and a sink-end B can be defined as:
  \begin{itemize}
    \item[-] $S(\Sigma)=\{q_0\}$, $I=\{A\}$, $O=\{B\}$, $s_0=q_0$
    \item[-] output function
      \begin{displaymath}
        \delta(\Sigma)(s,i)=\left\{
        \begin{array}[ht]{ll}
          (B:\bot) & s=q_0\land i=(A:\bot,B:\rnoread) \\
          (B:d) & s=q_0\land i=(A:d,B:\rread)\\
          \rblock & s=q_0\land i=(A:\bot,B:\rread) \\
          \rblock & s=q_0\land i=(A:d,B:\rnoread) \\
        \end{array}
        \right.
      \end{displaymath}
    \item[-] transition function $\lambda(\Sigma)(s,i)=q_0$.
  \end{itemize}
\end{example}

\begin{example}[PMM Semantics of n-Timer channel]
  Considering a n-Timer channel with a source-end A and a sink-end B, we define its PMM-based
  semantics as:
  \begin{itemize}
    \item[-] $S(\Sigma)=\{q_{i,d}|0\leq i\leq n, d\in \Sigma\}\cup\{q_0\}$
    \item[-] $I=\{A\}$, $O=\{B\}$, $s_0=q_0$
    \item[-] output function
      \begin{displaymath}
        \delta(\Sigma)(s,i)=\left\{
        \begin{array}[ht]{ll}
          (B:d) & s=q_{n,d}\land i=(A:d,B:\rread)\\
          \rblock & s=q_{n,d}\:\land \\
          & (i=T\lor i=(A:\_,B:\rnoread) \\
          \rblock & s=q_{j,d}\land 1 \leq j < n\land \\
          & (i=(A:\_,B:\rread)\lor \\
          & i=(A:d,B:\_))\\
          (B:\bot) & otherwise\\
        \end{array}
        \right.
      \end{displaymath}
    \item[-] transition function
      \begin{displaymath}
        \lambda(\Sigma)(s,i)=\left\{
        \begin{array}[ht]{ll}
          q_{j+1,d} & s=q_{j,d}\land i=T\land0 < j < n\\
          q_0 & s=q_{n,d}\land i=(A:\bot,B:\rread) \\
          q_{0,d} & s=q_{n,d'}\land i=(A:d,B:\rread) \\
          q_{0,d} & s=q_0 \land i=(A:d,B:\_) \\
          s & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
  \end{itemize}
\end{example}

@@

\todo{check the timed behavior}
\begin{definition}[Production Operator]
  Now we're going to define the production operator \emph{prod} of two parameterized Mealy machines as,
  \[
  prod(PM_1,PM_2)=PM_3
  \]
  as follows. Here we assume that $PM_2.O\cap PM_1.I=\varnothing$
  \begin{itemize}
  	\item[-] $\forall\Sigma, PM_3.S(\Sigma)=PM_1.S(\Sigma)\times PM_2.S(\Sigma)$
    \item[-] $PM_3.I=PM_1.I\cup PM_2.I-PM_1.O$
    \item[-] $PM_3.O=PM_1.O\cup PM_2.O-PM_2.I$
  \end{itemize}
  \emph{(Here we assume that sink ends of $PM_1$ can be connected to source ends $PM_2$, but not
  vise versa)}
  \begin{itemize}
    \item[-] $PM_3.s_0=(PM_1.s_0, PM_2.s_0)$
    \item[-] $\forall\Sigma, PM_3.\delta(\Sigma)((s_1,s_2), i)=$
      \begin{displaymath}
        \left\{
        \begin{array}[ht]{ll}
          (Out_1 + Out_2)|_{PM_3.O} & Out_1\neq\rblock\land Out_2\neq\rblock \\
          \rblock & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
      where we have
      \begin{itemize}
        \item[*] $In_1 = i|_{PM_1.I}$
        \item[*] $Out_1 = PM_1.\delta(\Sigma)(s_1,In_1)$
        \item[*] $In_2 = (Out_1 + i)|_{PM_2.I}$
        \item[*] $Out_2 = PM_2.\delta(\Sigma)(s_2,In_2)$
      \end{itemize}
    \item[-] $\forall\Sigma, PM_3. \lambda(\Sigma)((s_1,s_2),i)=(s_1',s_2')$
      where we have
      \begin{itemize}
        \item[*] $s_1' = PM_1.\lambda(\Sigma)(s_1,In_1)$
        \item[*] $s_2' = PM_2.\lambda(\Sigma)(s_2,In_2)$
      \end{itemize}
  \end{itemize}
\end{definition}

@@

  \begin{itemize}
  	\item[-] $\forall\Sigma, PM'.S(\Sigma)=PM.S(\Sigma)$
    \item[-] $PM'.I=PM.I-\{IN\}$
    \item[-] $PM'.O=PM.O-\{OUT\}$
    \item[-] $PM'.s=PM.s$
    \item[-] $\forall\Sigma, i\in Input(\Sigma,PM'.I,PM'.O)$, $PM'.\delta(\Sigma)(s, i)=$
      \begin{equation}
        \label{equ:existd}
        PM.\delta(\Sigma)(s,T)|_{PM'.O}
      \end{equation}
      if $i=T$. When $i$ is a non-time action, we need to check whether there's a data item
      $d\in\Sigma$
      satiefies
      \[
      PM.\delta(\Sigma)(s,i\cup\{IN:d\})|_{OUT} = d
      \]
      if \emph{true}, we have $PM'.\delta(\Sigma)(s, i)=$
      \[
      PM.\delta(\Sigma)(s,i\cup\{IN:d\})|_{PM'.O}
      \]
      otherwise $PM'.\delta(\Sigma)(s, i)=\rblock$
    \item[-] $\forall\Sigma, i\in Input(\Sigma,PM'.I,PM'.O).'$ the transition
      function can be defined as $PM.\lambda(\Sigma)(s,i)=$
      \begin{displaymath}
        \left\{
        \begin{array}[ht]{ll}
          PM.\lambda(\Sigma,T) & i=T \\
          PM.\lambda(\Sigma,i\cup\{IN:d\}) & \exists d\in\Sigma \mbox{ satisfes Equation
          \ref{equ:existd}} \\
          s & otherwise \\
        \end{array}
        \right.
      \end{displaymath}
  \end{itemize}
\end{definition}

@@

\section{From Blackbox to Timed Connectors} 
\label{sec:activelearning}
In this section, we will show how to use L* algorithm to extract models from blackboxes, which
mainly includes 3 steps: \emph{constructing hypothesis}, \emph{enclosing hypothesis} and
\emph{checking hypothesis}. Besides, we assume that input symbols and output symbols of the blackbox
have been provided as $\mathcal{I}$ and $\mathcal{O}$.

\subsection{Observation Table}
As mentioned above, membership queries provide us information with the form $mq(s)=o$, where $s$ is
an input sequence in $\Sigma^{+}$. Now the question is, \emph{how to find the connection between
such query results and Mealy machines?}

Generally, Mealy machines are composed of \emph{states} and \emph{transitions}.
Provided with a Mealy machine $m$, any input sequence in $m.I^+$ leads to a unique state. 
Such a sequence is called an \emph{access sequence} of this state.
Specially, we use $\varepsilon$, the \emph{empty sequnce}, to denote the initial state.
Similarly, if a blackbox model has the same bahavior with a Mealy machine, we can use access
sequences to label its states and, in the same way, input actions to label its transitions.
For example, we consider a Mealy machine stays is staying in some state labelled by $s$. With an
action $a$ provided, the Mealy machine will jump to a new state labelled by $s'=sa$ and generate a
output $mq(sa)$.

Different access sequences may lead to similar states. For example, a 2-Timer channel will dispense
any accepted data item in 2 time units. After that, the state of this channel is exactly same as its
initial state. Here \emph{suffixes} are used to distinguish different states as shown in the
following definition, where $D$ is initialized as $\mathcal{I}$. Under such a suffix set, two states
are considered different only when they have different one-step behaviors.

\begin{definition}[Equivalence under Suffixes]
  provided with two access sequences $s_1,s_2$ and a set of suffixes $D\subset\mathcal{I}^+$, we say
  the corresponding states of $s_1$ and $s_2$ are equivalent under $D$ (denoted as $s_1\sim_D s_2$)
  iff $\forall d\in D, mq(s_1d) = mq(s_2d)$.
\end{definition}

Since all this \emph{inputs} and \emph{outputs} can be observed outside the
model, we can use the same notations to describe a black box as a hypothetical Mealy machine.
Following this idea, \emph{Observation Tables}, proposed in \cite{DBLP:journals/iandc/Angluin87},
shows a tabular form of such hypothetical models. In observation tables, rows represent states with
label of their access sequences and columns are labelled suffixes. Successors of a state $s$ is,
intuitively, denoted as extended access sequences $sa$ where $a$ is a single input action. 
A cell with column-label $d$ and row-label $s$ should contain the value $mq(sd)$. Algorithm
\ref{alg:buildtable} shows how to build such an observation table.

\begin{algorithm} 
  \caption{BuildTable} 
  \label{alg:buildtable}
  \KwIn{Oracle interface $mq$, Input actions $\mathcal{I}$, suffix set $D$}
  \KwOut{Observation table $obs$} 
  $obs$ initialized as empty\;
  $unclosed=\{\varepsilon\}$\;
  \Repeat{$unclosed=\varnothing$}
  {
    $next=\{st|\forall s\in unclosed,\forall t\in \mathcal{I}\}$\;
    append $unclosed$ to $obs$\;
    $unclosed=\{seq\in next|\forall a\in obs, seq \not\sim_D a\}$\;
  }
  \Return $obs$\; 
\end{algorithm}

Taking a 2-Timer channel as an example, we now illustrate how this algorithm works. 
We assume that the source end of this 2-Timer channel is $A$, the sink end is $B$ and the alphabet
is $\{a\}$. We briefly denote $\{A:a,B:\rnoread\}$ and other input/output actions in form of
$a,\rnoread$.

Firstly, $obs$ is initialized as empty and the empty sequence $\varepsilon$ is pushed in $unclosed$.
Then we will explore all the access sequences in $unclosed$ and calculate its successors. Here the
$unclosed$ set consists of access sequences that has no equivalent fellow with less length. 
For example, figure \ref{fig:hypo} shows the $obs$ of a 2-Timer channel after the first
iteration. The five successors of $\varepsilon$ is presented as the bottom part of the table, where
four are equivalent with $\varepsilon$ but $a,\rnoread$ is not. Therefore, we take $a,\rnoread$ as a
brand-new state and all of its successors need further exploration (see in the hypothesis Mealy
machine presented in the following figure).

\begin{figure}[ht]
  \begin{center}
    \begin{displaymath}
      \begin{array}{l||ccccc}
        \hline
        & a,\rread & \bot,\rread & a,\rnoread & \bot,\rnoread & T\\
        \hline\hline
        \varepsilon & \rblock & \rblock & \rempty & \rempty & \rempty \\
        \hline
        a,\rread & \rblock & \rblock & \rempty & \rempty & \rempty \\
        \bot,\rread & \rblock & \rblock & \rempty & \rempty & \rempty \\
        \rowcolor[gray]{.9}
        a,\rnoread & \rblock & \rblock & \rblock & \rempty & \rempty \\
        \bot,\rnoread & \rblock & \rblock & \rempty & \rempty & \rempty  \\
        T & \rblock & \rblock & \rempty & \rempty & \rempty \\
        \hline
      \end{array}
    \end{displaymath}
    \input{./images/hypo.tex}
  \end{center}
  \caption{Observation Table and Corresponding Hypothesis}
  \label{fig:hypo}
\end{figure}

In the second iteration, we explore the successors of $a,\rnoread$. Fortunately we find that every
successor has a shorted equivalence fellow (\emph{itself}). Consequently, the algorithm terminates
with a closed hypothesis where all the unclosed edge shown in Figure \ref{fig:hypo} turn into
self-loop.

\begin{figure}[ht]
  \begin{center}
    \begin{displaymath}
      \begin{array}{l||ccccc}
        \hline
        & a,\rread & \bot,\rread & a,\rnoread & \bot,\rnoread & T\\
        \hline\hline
        \varepsilon & \rblock & \rblock & \rempty & \rempty & \rempty \\
        \hline
        a,\rread & \rblock & \rblock & \rempty & \rempty & \rempty \\
        \bot,\rread & \rblock & \rblock & \rempty & \rempty & \rempty \\
        a,\rnoread & \rblock & \rblock & \rblock & \rempty & \rempty \\
        \bot,\rnoread & \rblock & \rblock & \rempty & \rempty & \rempty  \\
        T & \rblock & \rblock & \rempty & \rempty & \rempty \\
        \hline
        a,\rnoread-a,\rread & \rblock & \rblock & \rblock & \rempty & \rempty \\
        a,\rnoread-\bot,\rread & \rblock & \rblock & \rblock & \rempty & \rempty \\
        a,\rnoread-a,\rnoread & \rblock & \rblock & \rblock & \rempty & \rempty \\
        a,\rnoread-\bot,\rnoread & \rblock & \rblock & \rblock & \rempty & \rempty \\
        a,\rnoread-T & \rblock & \rblock & \rblock & \rempty & \rempty \\
        \hline
      \end{array}
    \end{displaymath}
  \end{center}
  \caption{A Closed Observation Table}
  \label{fig:hypo2}
\end{figure}

\subsection{Counter-Examples' Analysis} 
In our framework, the oracle function $eq$ is used to check the equivalence between hypothesis
models and blackboxes. It is also responsible for generating counter-examples.
Generally, as mentioned in \cite{DBLP:journals/iandc/Angluin87,DBLP:conf/sfm/SteffenHM11},
equivalence query has been proved impossible in blackbox models.

Apparently, the closed hypothesis presented in Figure \ref{fig:hypo2} is not equal to the 2-Timer
channel. It's easy to find a counter-example $s=a,\rnoread-T-T-\bot,\rread$ where $mq(s)=a$ while
according to the hypothesis, the result shoud have been $\rempty$. As mentioned above, we use a
suffix set $D$ to distinguish different states. The existance of counter-examples shows that the
current suffix set $D$ is not powerful enough to distinguish all different states. We need to analyze
the counter-example and in turn append more elements to $D$.

@@

\subsection{Case Studies}
A simple example of timed connector is presented to show how L* works.
\begin{figure}[ht]
  \begin{center}
    \input{./images/expfifo.tex}
  \end{center}
  \caption{Expiring FIFO1 Channel (ExpFIFO1)}
  \label{fig:expfifo}
\end{figure}

@@

\begin{table}[ht]
  \renewcommand{\arraystretch}{1.3}
  \caption{Time-Cost Analysis}
  \label{tabel:timecost}
  \centering
  \begin{tabular}{l||rrrr}
    \hline
    & FIFO & Alternator & Gate & ExpFIFO1\\
    \hline\hline
    Membership Query(s) & 41.571 & 126.468 & 169.161 &\\
    Hypothesis Query(s) & 0.001 & 0.003 & 0.004 &\\
    Total Time(s) & 41.715 & 165.114 & 247.098 &\\
    Membership Query(\%) & 99.6 & 76.6 & 68.5 &\\
    \hline
  \end{tabular}
\end{table}

@@

If previous simulation results are stored in a well-maintained cache, the time-cost in
simulation process could be reduced signficiantly. In this work, we use a multiway tree to buffer
these results.

\begin{table}[ht]
  \renewcommand{\arraystretch}{1.3}
  \caption{Reduction of Membership Queries}
  \label{tabel:cacheoptimization}
  \centering
  \begin{tabular}{l||rrr}
    \hline
    & FIFO & Alternator & Gate \\
    \hline\hline
    Original Algorithm & 93 & 880 & 1034 \\
    Cached Algorithm & 90 & 725 & 707 \\
    Reduction Rate & 3.2\% & 21.4\% & 31.6\% \\
    \hline
  \end{tabular}
\end{table}

With cache applied, we have made considerable reduction on the calls of membership queries. The results
can be found in Table \ref{tabel:cacheoptimization}.
